# Introduction: R Basics

For the love of God before you do anything, familiarize yourself with R Projects and the `here` package. These make R so much more user friendly and less of a nightmare. If you need an overview, go here: http://jenrichmond.rbind.io/post/how-to-use-the-here-package/


Now lets get stuck in.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


## Importing Data

### Spreadsheets

See https://nacnudus.github.io/spreadsheet-munging-strategies/index.html for more detailed and in-depth tutorials (if you need that kind of thing)


## Exporting (i.e., saving) Data and Output

### Exporting to .CSV

Generally speaking, unless you have a specific reason to, don't. But if you must: `write_csv()`

### Export to .RData (and load the data again later)

```{r eval=FALSE}
save(obj_name, file=here::here("subfolder", "save_file_name"), compress = FALSE)

load(here::here("folder", "save_name.RData"))
```


### Export to Excel

```{r}
library(openxlsx)

#Method 1: If you only want to export 1 thing, and/or only need output document
  #write as object, with no formatting:
  write.xlsx(objectname,file = "filenamehere.xlsx",colnames=TRUE, borders="columns")

  #write as table:
  write.xlsx(objectname,"filename.xlsx",asTable = TRUE)


#Method 2: If you want to do the above, but add multiple objects or tables to one workbook/file:
  ## first Create Workbook object
  wb <- createWorkbook("AuthorName")
  #then add worksheets (as many as desired)
  addWorksheet(wb, "worksheetnamehere")
  
#then write the object to the worksheet  
writeData(wb, "test", nameofobjectordataframe, startCol = 2, startRow = 3, rowNames = TRUE)

#save excel file
saveWorkbook(wb, "filenamehere.xlsx", overwrite =TRUE)


#Method 3: exact same as method 2, but creating a more fancy tables

wb <- createWorkbook("Ryan")
addWorksheet(wb, "worksheetnamehere")
writeDataTable(wb, sheetName, objectName, startCol = 1, startRow = 1, colNames = TRUE, rowNames = FALSE,
          tableStyle="TableStyleLight2",tableName=NULL, headerStyle = NULL,withFilter=FALSE,keepNA=TRUE,sep=", ",
          stack = FALSE, firstColumn = FALSE, lastColumn = FALSE,bandedRows = TRUE,bandedCols = FALSE)

saveWorkbook(wb, "filenamehere.xlsx", overwrite =TRUE)

```


### Access/edit specific cell number values

```{r}
rainbow=tibble::tribble(~Color,
                "red",
                "orange",
                "black",
                "green",
                "blue",
                "purple")

rainbow$Color[3] # access, but can't overwrite this way
rainbow[3,"Color"] # access and can overwrite

rainbow[3, "Color"]= "yellow" # save this value to row 3 in column "Color"

rainbow
```




