# Wrangle Data

This chapter contains useful tips on wrangling (i.e., manipulating) data. If you need to know to do to things like create new variables, split one variable into multiple variables, pivot a data set from wide to long, etc., look no further.

If you want a pretty good intro tutorial to the `dplyr` package, click [here](https://www-r--bloggers-com.cdn.ampproject.org/v/s/www.r-bloggers.com/2021/01/how-to-analyze-data-with-r-a-complete-beginner-guide-to-dplyr/amp/?amp_js_v=a6&gsa=1&usqp=mq331AQFKAGwASA%3D#csi=0&referrer=https%3A%2F%2Fwww.google.com&tf=From%20%251%24s&ampshare=https%3A%2F%2Fwww.r-bloggers.com%2F2021%2F01%2Fhow-to-analyze-data-with-r-a-complete-beginner-guide-to-dplyr%2F)


## Select helpers

Sometimes when working with a data set, you want to work with a few *specific* variables. For instance, maybe you want to view a graph of only reverse-coded variables (which start with the prefix "r"); or maybe you want to create a subset of your data that has a few specific variables removed. For this you can use `dplyr::select()` and its associated helper commands

`select()` can be thought of as "extract"; it tells R to identify and "extract" a specific variable (or variables)

```{r 3.8, eval=FALSE}
cars=mtcars

# select one column
cars %>% select(mpg)

# select multiple columns, if they are all next to one another
cars %>% select(mpg:hp)

# select multiple columns by name (when not next to one another) by defining them in a vector
cars %>% select(c(mpg, hp, wt))

# select only variables that start with a certain prefix/character/pattern/etc.
cars %>% select(starts_with("d"))

# ...or columns that end with a certain prefix/etc.
cars %>% select(ends_with("t"))

# ...or contains a certain pattern or string
cars %>% select(contains("se"))

# select ALL OF the variables in a data set that match those of a pre-defined vector
  
  # first define the names in a vector
  vars=c("hp", "drat", "gear", "carb")
  
  #now use helper
  cars %>% select(all_of(vars))
  
# select ANY OF the variables in a pre-defined vector
  
  vars_2=c("hp", "drat", "watermelon", "grilled_cheese") # only the first two will be in the data
  
  cars %>% select(any_of(vars_2)) # only (and all of) the variables actually PRESENT in the data are pulled
  
# select only variables of a certain class or type
  cars %>% select(where(is.numeric))
  cars %>% select(where(is.character))
```

Other examples can be seen on [THIS LINK](https://tidyselect.r-lib.org/reference/language.html) for a simple but detailed guide.

## Advanced Filtering techniques

All info here and the code taken from [this link](https://suzan.rbind.io/2018/02/dplyr-tutorial-3/#filtering-based-on-a-exact-character-variable-matches); credit goes to Suzan Baert.

### Filter based on partial match

Combine `filter` with `str_detect` to search for pattern matches in a column

```{r}

```

### Filtering based on multiple conditions

You can filter for more than one value at once, or combine operators to do searches with conditions....

| Goal                                                                                     | Code                                 |
|------------------------------------------------------------------------------------------|--------------------------------------|
| Return only rows where ***both*** conditions are met                                     | `filter(condition1, condition2)`     |
| Return all rows where condition 1 is true but condition 2 is not                         | `filter(condition1, !condition2)`    |
| Return rows where condition 1 ***and/or*** condition 2 is met                            | `filter(condition1 | condition2)`    |
| Return all rows where only one of the conditions is met, ***and not when both are met*** | `filter(xor(condition1, condition2)` |

The sample code will return all rows with a bodywt above 100 and either have a `sleep_total` above 15 or are not part of the Carnivora order.

```{r advanced_filter_1, eval=FALSE}
msleep %>% 
  select(name, order, sleep_total:bodywt) %>% 
  filter(bodywt > 100, (sleep_total > 15 | order != "Carnivora"))
```

### Filtering across multiple columns

From Susan Baert's blog page:

You have three options for how to do this:

-   `filter_all()`, which filters columns *based on your further instructions*

-   `filter_if()`, which requires a boolean to indicate which columns to filter on. If that is true, the filter instructions will be followed.

-   `filter_at()`, which requires you to specify columns inside a `vars` argument for which the filtering will be done.

Retain all rows with the pattern match "Ca" inside. Useful for when you want to search for a key pattern (or number) across multiple columns.

```{r advanced_filter_2, eval=FALSE}
msleep %>% 
  select(name:order, sleep_total, -vore) %>% 
  filter_all(any_vars(str_detect(., pattern = "Ca")))

msleep %>%  
  select(name, sleep_total:bodywt) %>% 
  filter_all(any_vars(. < 0.1))

# You can also switch any_vars to all_vars to filter across the whole data frame
msleep %>%  
  select(name, sleep_total:bodywt, -awake) %>% 
  filter_all(all_vars(. > 1))
```

What if you want to filter specific column types in your data frame, like just date columns, to find a specific date across multiple date columns? Or filter across multiple number columns to find every instance of a specific number? `filter_if` is better than `filter_all` here because the latter would return a filter search across the whole data frame. By using `filter_if`, we can get more specific about which columns to search through.

```{r advanced_filter_3, eval=FALSE}
msleep %>% 
  select(name:order, sleep_total:sleep_rem) %>% 
  filter_if(is.character, any_vars(is.na(.)))
```

The last command, `filter_at`, does things differently. It doesn't filter all columns at once; and you do not have to specify the column type or which columns. ***This one allows you to indicate which columns to search through like you would within any select statement***.

Example: Search through the columns that start with "sleep", for values in all of those columns above .5.

```{r advanced_filter_4, eval=FALSE}
msleep %>% 
  select(name, sleep_total:sleep_rem, brainwt:bodywt) %>% 
  filter_at(vars(contains("sleep")), all_vars(.>5))
```

## If-then and Case-when

The premise of an if/then or if/else statement is simple: "If condition 1 is satisfied, perform x operation; if not, then do y"

```{r 3.9}
mtcars %>% mutate(power_level=ifelse(mtcars$hp<350, "Low", "High")) %>% head()
```

This line of code effectively says: if the length in Sepal.Length is \>5, set new variable = to "short"; else, set it to "long"

Case-when

When you have 3+ conditions, it's easier to use case-when. This is a more simple and straightforward approach than nesting multiple if-else commands

```{r 3.10, eval=FALSE}
My_vector= case_when(
	Condition1 ~ value1,
	Condition2 ~ value2,
	Condition3 ~ value3
	TRUE ~ valueForEverythingElse #catch all for things that don't meet the above conditions
	)
```

Example:

```{r 3.11}

mtcars %>% mutate(size= case_when(cyl==4 ~ "small",
                                  cyl==6 ~ "medium",
                                  cyl==8 ~ "large")) %>% 
  select(c(cyl,size)) %>% head()
```

## Dataframe Manipulation

### Pivoting (i.e., transposing) data

Pivot wide to long

Rearranging data like this can make it easier to work with and analyze. Example below from my gradebook for stats (exported from Canvas), with fake names.

The command structure is as follows:

```{r 3.16, eval=FALSE}

 pivot_longer( # Transpose LENGTHWISE by....
    cols = everything(), # Taking ALL variable names...
    names_to="variable", # ...and dumping them into this new variable/column
    values_to="missing_count") #...and placing their values in this other new column
```

**NOTE!!!** Pivoting data from wide to long like this expands the number of rows to make a matrix so that (for example, each student now has as a row for each assignment). Therefore, you can only pivot longways (or wide) **ONCE**, otherwise you will make duplicates.

*If you need to pivot multiple columns, just include all of the columns in one single pivot; do not use two separate, back to back pivot commands.*

Example:

```{r 3.17}
gradebook=tibble::tribble(
  ~Student, ~Homework.1, ~Homework.2, ~Homework.3, ~Homework.4, ~Homework.5, ~Quiz.1, ~Quiz.2, ~Quiz.3, ~Quiz.4, ~Final,
     "Bob",         19L,          0L,          13,          16,          0L,      21,      7L,      15,    17.5,     33,
    "Jane",         17L,         19L,          16,        16.5,         25L,    21.5,     19L,   14.75,     9.5,   39.5,
    "John",         19L,         19L,        14.5,        19.5,         25L,      21,     21L,    18.5,      17,   46.5
  )

head(gradebook)

gradebook=gradebook %>% 
   pivot_longer( # Transpose lengthwise by:
    cols = Homework.1:Final, # Taking these variables
    names_to="Assignment", # ...and dumping them into this new variable, storing them lengthwise
    values_to="Points") #...then place their values in this new column

gradebook %>% head()
```

### Turn Row Names into a Variable

Use the `rownames()` command to turn row names into a variable

```{r}
cars=rownames_to_column(mtcars, var = "car")

as_tibble(cars) %>% slice(1:6)
```

### Rename Variables

TWO WAYS TO DO THIS: Use `colnames()` (for base R) or `rename()` (for tidyverse)

`colnames()` pulls up all the column/variable names as a vector. If you want to actually change them, you'll need to combine this command with something like the sub() or gsub() commands (for base R). I'm going to skip this because...its base R.

To access and change the names faster via tidyverse, run use `rename()`

```{r}
rm(list=ls()) # clear R's memory

iris %>% rename("hurr"="Sepal.Length",
                "durr"="Sepal.Width",
                "abcdefgh"="Species") %>% 
  head()
```

#### Rename multiple variables at once
If you need to do some really fancy conditional renaming (e.g., changing all variables that start with "r" to start with "rf" instead, to make it more clear that the prefix actually stands for "risk factor" rather than "reverse coded"), you'll need to use `rename_with()`.

This command has two parts to it: the data set, and the function you wish to apply to it (which you put after the `~`)

```{r }
rename_with(iris, ~ gsub(pattern = ".", replacement = "_", .x, fixed = TRUE)) %>% 
  head()
```

The gsub() function from Base R identifies matching patterns in the data and substitutes them with what you want instead. Think of it like R's version of Find/Replace from Microsoft Word.

The above line of code thus does the following: 1. First, it checks the column names of the supplied data set (`iris`) for a specific pattern (specified in `pattern=` ) 2. Then it replaces that pattern with your input in `replacement=`

The great thing about rename_with() is that the .fn (or `~` for short) can take *ANY* function as input. For example, if you want to **add** an element to the column names rather than replace something, (e.g., a prefix or suffix), you can change the function to:

```{r}
rename_with( iris, ~ paste0(.x,  "_text")) %>% 
  head()
```

The above line adds a suffix. You can also add a prefix in the exact same way, just by switching the order of the string and the pattern in the paste0 command.

***Alternative method to the above*** This is a second way to do the above. It may appear more simple, but it's also probably not as theoretically consistent with how the packages were made.....it uses the `stringr` package to rename the column names, and `stringr` is typically used for editing vectors of strings in a data set.

...so it works, but it's a little unconventional because you call and edit the column names like you would a variable in your data set.

```{r}
colnames(iris)=str_replace(colnames(iris), pattern = ".", replacement = "_")
```

In short: `rename()` and `rename_with()` are for *renaming* variables, as their names imply. The `str_` verbs from the stringr package are for editing string-based variabels in your data set. Either works though with a little ingenuity.



### Change column order in a data set

Use `relocate()` to change column positions. If you need to move multiple columns at once, this command uses the same syntax as `select()`.

```{r}

mtcars # notice the column order

mtcars %>% relocate(hp:wt, .after= am) %>% head()
```

Alternatively, you could use this solution:
```{r}
# Assuming your variables are in a data frame called 'df'
# Get names of variables starting with "r"
r_vars <- names(df)[grepl("^r", names(df))] 

# Create new names
new_names <- paste0("item_", sub("^r", "", r_vars))

# Rename the variables
df <- df %>% rename_with(~ new_names[match(., r_vars)], .cols = all_of(r_vars))
```


### Splice/Join Dataframes

Joining and splitting data is pretty straightforward....

### Whole Data Sets

The code below is from [this excellent tutorial](https://www.youtube.com/watch?v=SCdmyyoudb8&t=23s)

```{r chapter_3}
set.seed(2018)

df1=data.frame(customer_id=c(1:10),
               product=sample(c('toaster','TV','Dishwasher'),10,replace = TRUE))


df2=data.frame(customer_id=c(sample(df1$customer_id, 5)),state=sample(c('New York','California'),5,replace = TRUE))

df1=tibble::as_tibble(df1)
df2=tibble::as_tibble(df2)

# df1 =left table
# df2= right table
```

Inner join - retains only rows with values that appear in both tables, and matches by keys.
Left join - returns everything in the left, and rows with matching keys in the right
Right join - returns everything in the right, and rows with matching keys in the left
Full join - retain all rows from both tables, and join matching keys in both right and left
Anti join - returns all rows in the left that do not have matching keys in the right

*If you're joining two Qualtrics surveys together, this is most likely the one you want to use (e.g. matching by participant name, and only keeping rows in the joined data set for participants that have responses logged in both survey 1 and survey 2*

```{r 3.2}
df1 %>% inner_join(df2,by='customer_id')
df1 %>% left_join(df2,by='customer_id')
df1 %>% right_join(df2,by='customer_id')
df1 %>% full_join(df2,by='customer_id')
df1 %>% anti_join(df2,by='customer_id')
```


## Working with Dates/Times

Formatting a column of dates can be extremely helpful if you need to work with time data, but also an extreme pain in the ass if it's not stored correctly. This tutorial will be divided into two parts to cover both scenarios that you could encounter. It requires things to be done in two stages, and very precisely.

### Date-time objects

If you're lucky enough to have a vector of date-times, like what Qualtrics gives you, this will be brainless. Just do the following:

```{r}
example_datetime_data=tibble::tribble(~datetime,
                                      "2010-08-03 00:50:50",
                                      "2010-08-04 01:40:50",
                                      "2010-08-07 21:50:50")

head(example_datetime_data) # stored as character string

# Tidyverse
lubridate::as_date(example_datetime_data$datetime)
```

### Date-only objects

If you're unlucky enough to have *only* dates, and said dates are written in the traditional **x/x/xxxx** format, this will be an annoyance that has to be done in two stages.

First, assuming your data is already imported and is being stored as a vector of character strings, you have to tell R to adjust the formatting of dates. You cannot change it from a character-based object into a Date or DateTime one until it recognizes the correct formatting.

```{r}
example_date_data=tibble::tribble(~X1,     ~X2,
                                  "8/4/2021",  -49.87,
                                  "8/4/2021",  -13.85,
                                  "8/3/2021",   -7.45,
                                  "8/3/2021", -172.71)

# Correct formatting
example_date_data$X1=format(as.POSIXct(example_date_data$X1,format='%m/%d/%Y'),format='%Y-%m-%d')

head(as_tibble(example_date_data))
```

In the code above, note that there are two `format` commands: The first one tells R how the date data is **currently being stored**, while the second at the end tells it how **you want it to be stored.** In this case, we are changing it from the way we would usually hand write a date (e.g., 10/26/1993) to a format commonly recognized and used in Excel and stats software (1993-10-26). *If your column also has times in it, you also need to include that too!*

Second, you can now correct the object's structure. You can do this with base R's `as.Date()` or tidyverse's `date()` verbs.

```{r eval=FALSE}

# Tidyverse
example_date_data$X1= lubridate::date(example_date_data$X1) 

# Base R version
example_date_data$X1=as.Date(example_date_data$X1)
```

Notice how the object is now stored as the correct type in the table above.

***NOTE!*** This entire process has been included in the `tidy_date()` command in my package, `legaldmlab`.

### Find the difference between two dates/times

```{r , eval=FALSE}
difftime(part_1$end_date[1], part_2$end_date[1], units="days")
```


## Variable Coding and Creation

### Reverse-code a variable

To reverse-score a variable, you should use `car::recode()` Can be done a few different ways, depending on how many variables you're looking to recode:

```{r eval=FALSE}

# Recode just one variable
df$column=recode(df$column,"1 = 7 ; 2 = 6 ; 3 = 5 ; 5 = 3 ; 6 = 2 ; 7 = 1")

# Recode a select bunch of variables
df=df %>% mutate(across(c(family_close : family_feelings), recode, "1 = 7 ; 2 = 6 ; 3 = 5 ; 5 = 3 ; 6 = 2 ; 7 = 1"))
```

### Dummy coding (the very fast and easy way)

Use dplyr's `pivot_wider` in conjunction with `mutate` to very quickly and automatically dummy code a column with any number of unique values. The middle part of the code below is what you need...just copy and paste it, and tweak the specifics

```{r}
library(tidyverse)

mtcars |> 
  mutate(car=rownames(mtcars)) |> 
  
  dplyr::mutate(n=1) |> 
  tidyr::pivot_wider(names_from = cyl,
              values_from = n,
              names_prefix = "number_cyl",
              values_fill = list(n=0)) |> 
  
  select(car, starts_with("number_")) |> head() #truncate output for easier reading
```

### Create a relative ranking variable

If you want to create a variable that is an ordinal ranking of other variables, first you need to make sure your data is long-wise. Then, depending on the type of ranking system you want, you'll might need a different ranking command....

-   The `min_rank` command from dplyr works in a manner similar to base R's `rank` command. It ranks things like you see in sporting events. For example, if there is a clear winner in a game but 3 people tie for second place, the ranks would look like this: 1,2,2,2,4,5. Notice that the positions are independent from the counts.

-   Using the same example from above, if you want the ranks to have no gaps (i.e. 1,2,2,2,3,4), you need to use dplyr's `dense_rank` command.

-   In either case, the ranks are generated from lowest to highest, so if you want to flip them around you'll need to include `desc()` in the command.

```{r chapter_3_end}

dat=tibble::tribble(~name, ~score,
                    "bob", 0,
                    "bob", 5,
                    "bob", 50,
                    "bob", 50,
                    "bob", 50,
                    "bob", NA,
                    "alice", 70,
                    "alice", 80,
                    "alice", 90,
                    "alice", 20,
                    "alice", 20,
                    "alice", 1)

dat %>% mutate(ranked = dense_rank(desc(score)))
```

### Merging variables

Sometimes you'll have multiple variables and you want to collapse them into a single variable. The `pmin()` command is useful for this.

``` {r3.14}
example_data=tribble(~A,~B,~C,
                     1,NA,NA,
                     2,NA,NA,
                     3,NA,NA,
                     NA,4,NA,
                     NA,5,NA,
                     NA,6,NA,
                     NA,NA,7,
                     NA,NA,8,
                     NA,NA,9)

example_data %>%
  mutate(accept_reject = 
           pmin(A,B,C,na.rm = TRUE))
```


### Mutate multiple variables at once

You can either specify each column individually, like above, or tell R to identify columns for you based on their type or their name. This requires adding in one additional verb--either contains() or where() depending on what you want to do.

Two simple examples:

```{r 3.15}
# turn multiple variables into factors
ex_data=dplyr::tribble(~color, ~car,
                       "red", "corvette",
                       "blue", "chevelle",
                       "green", "camaro",
                       "red", "corvette",
                       "green", "chevelle",
                       "yellow", "gto")

dplyr::glimpse(ex_data)

ex_data %>% mutate(across(c(color, car),factor))

# round multiple columns to 1 decimal place
mtcars %>% mutate(across(c(disp:qsec),round,1)) %>% head()
```


### Conditional replacement of values

The following code is useful if you want to replace a value in one column, and the replacement is conditional upon the value in another column.

```{r 3.12}
mpg %>% 
  mutate(across(.cols = c(displ, cty, hwy),
                .fns = ~case_when(cyl == 4L ~ as.numeric(NA),
                                  TRUE ~ as.numeric(.x))))
```

```{r 3.13, eval=FALSE}
test %>% 
  mutate(across(.cols = c(rank),
                .fns = ~case_when(is.na(participant_score) ~ as.numeric(NA),
                                  TRUE ~ as.numeric(.x))))
```


## The working environment

### Stuff the WHOLE working environment into a list

```{r eval=FALSE}
files=mget(ls())
```

### Extract everything from a list into the environment

```{r eval=FALSE}
list2env(cog_data, globalenv())
```

### Delete everything in the entire environment, except for one item

```{r eval=FALSE}
rm(list=setdiff(ls(), "cog_data")) # delete everything in the local environment except the final data set
```

## Lists

### Nesting

Imagine the concept of Russian Dolls, applied to data sets. You can manage data sets more effectively my collapsing them into a single tiny, mini data frame, and stuffing that inside of another one. This is done via "nesting"...

Effectively, you smush/collapse everything down so it fits inside one column. You can unnest to expand this data back out later when you need it, and keep it collapsed when you don't. This works because a vector/column in a data frame *is* a list of a defined length; and a data frame is thus simply a collection of lists that are all the same length.

-   You can store anything in a data frame. You can keep the df connected to the model, which makes it very easy to manage a whole slew of related models

-   You can use functional programming (i.e., iterative functions) to map functions or combinations of functions in new ways. Moreover and more importantly, when you use `purrr` to map functions onto multiple models or objects simultaneously, you're doing it to all of them at once with a single command, and the objects are kept together while you do it. This limits the mistakes you can make (e.g., copying and pasting code and forgetting to tweak something important; applying a function to the wrong object or set of objects by accident), and also reduces unnecessary code in your script.

-   Converting data into tidy data sets gives you a whole new way (and easier way) to manage lots of information

```{r 3_nesting}
head(mtcars |> nest(crap=vs:carb))
```

### Combining/collapsing list levels (Reducing)

```{r eval=FALSE}
demo_vars=files |> 
  map(import_spss) |> 
  reduce(left_join, by="catieid")
```

## String Variables

### Remove a pattern from a string

```{r chapter_6}

price_table=tribble(~car, ~price,
        "Corvette", "$65,000",
        "Mustang GT", "$40,000")

# BASE R METHOD (sub by replacing something with nothing)
gsub("\\$", "",price_table$price) # (pattern, replace with, object$column)

# TIDYVERSE METHOD
str_remove(price_table$price, pattern = "\\$")
```

You can remove numbers by typing `"[:digit:]"`
```{r 6.1, eval=FALSE}
panss_sem_data$cgi_sev=str_remove(panss_sem_data$cgi_sev, pattern = "[:digit:]")
```


### Replace one pattern in a string with another

Tidyverse command: `str_replace()` or `str_replace_all()`
Base R command: `gsub()`

```{r 6.2, eval=FALSE}
# base R
gsub(mtcars, replacement = )

#tidyverse
str_replace_all(iris$Species, pattern=c("e", "a"), replacement="ZZZZ") |> 
  head()

str_replace(iris$Species, pattern=c("e", "a"), replacement="ZZZZ") |> 
  head()

```


### Find (i.e., filter for) all instances of a string

Useful for finding very specific things inside a column (e.g., one particular person's name in a roster of names; everyone with a particular last name)

Tidyverse command: `str_detect()`
Base R command: `grepl()`

Note both must be nested inside of `filter()`
```{r 6.3, eval=FALSE}

cars_df=rownames_to_column(mtcars, var = "car")

# base R
cars_df |> filter(grepl("Firebird", car))

# tidyverse
cars_df %>% filter(str_detect(car,"Firebird"))
```

You can also search for **multiple strings simultaneously** by including the "or" logical operator *inside* the quotes.

```{r 6_filtermultiple, eval=FALSE}
cars_df |> filter(str_detect(car, "Firebird|Fiat"))
```

You can also include the negation logical operator to filter for all instances *except* those with the specified string.

```{r eval=FALSE}
# base R
cars_df |> filter(!(grepl("Pontiac", car)))

# tidyverse
cars_df |> filter(!(str_detect(car, "Pontiac")))
```

### Drop all rows from a data set that contain a certain string

```{r eval=FALSE}
# Tidyverse method
cars_df |> 
  filter(str_detect(car, "Merc", negate = TRUE)) #including negate=TRUE will negate all rows with the matched string

# base R
cars_df[!grepl("Merc", cars_df$car),]
```



### Force all letters to lower case

Use `stringr::str_to_lower()`

```{r 6.5, chapter_6_end}
blah=tribble(~A, ~B,
             "A","X",
             "A","X")

blah

blah$A=str_to_lower(blah$A)

blah
```


### Replace strings by via keywords

```{r}

lookup_table=tribble(
  ~keyword,                         ~mapping,
  "trello.com* atlassian",          "services",
  "microsoft*microsoft 365 p",      "services",
  "expressvpn",                     "services",
  "1password",                      "services",
  "fing",                           "services",
  "gsuite",                         "services",
  "flightradar",                    "services",
  "amazon prime",                   "services",
  "trailforks",                     "services",
  "alltrails",                      "services",
  "grammarly",                      "services",
  
  "elderscrollsonline",             "entertainment",
  "steam",                          "entertainment",
  "microsoft*xbox",                 "entertainment",
  "qobuz",                          "entertainment",
  "amazon music",                   "entertainment",
  "disney mobile",                  "entertainment",
  "espn",                           "entertainment",
  "audible",                        "entertainment",
  "kindle svcs",                    "entertainment",
  "rock ridge golf",                "entertainment",
  "prime video",                    "entertainment",
  "coyote creek",                   "entertainment",
  "archery headquarters",           "entertainment",
  "microsoft*xbox",                 "entertainment",
  
  "shoprite",                       "groceries",
  "aldi",                           "groceries",
  "target",                         "groceries",
  "trade",                          "groceries",
  "alma coffee",                    "groceries",
  "krauzers",                       "groceries",
  "apollo liquor",                  "groceries",
  
  "salon",                          "health",
  "pharmacy",                       "health",
  "mayo",                           "health",
  "mychart",                        "health",
  "walgreens",                      "health",
  "cvs",                            "health",
  "amznpharma ",                    "health",
  "nystrom",                        "health",
  
  "ebay",                           "shopping",
  "worldwide cyclery",              "shopping",
  "ua com",                         "shopping",
  "wal-mart",                       "shopping",
  "walmart",                        "shopping",
  "wm supercenter",                 "shopping",
  "bestbuy",                        "shopping",
  "martys reliable",                "shopping",
  "kali",                           "shopping",
  "shop nhl",                       "shopping",
  "apple",                          "shopping",
  "amazon mktp",                    "shopping",
  "amazon com",                     "shopping",
  "amazon mark",                    "shopping",
  "amazon ret",                     "shopping",
  "amazon digit",                   "shopping",
  "amzn mktp",                      "shopping",
  "atm",                            "shopping",
  "gamersnexus",                    "shopping",
  "drop com",                       "shopping",
  "reebok",                         "shopping",
  "gold tip",                       "shopping",
  "hacklebarney farm",              "shopping",
  
  "wendy",                          "dining",
  "taco bell",                      "dining",
  "moe s",                          "dining",
  "subway",                         "dining",
  "domino",                         "dining",
  "starbucks",                      "dining",
  "coffee potter",                  "dining",
  "burnt mills cider",              "dining",
  "jersey girl brewin",             "dining",
  
  "parkmobile",                     "travel",
  "uber   trip",                    "travel",
  
  "withdrawal from dept education student ln", "debt",
  "withdrawal from kmf kmfusa.com",             "debt"
)

# Correctly map categories by keywords from table
spending <- spending |> 
  mutate(keyword = str_extract(description, paste(lookup_table$keyword, collapse = "|"))) |> 
  left_join(lookup_table, by = "keyword") |> 
  mutate(category = coalesce(mapping, category)) |> 
  select(-keyword, -mapping)

```

## Factors

### Manually recode/change a factor's levels

Use `forcats::fct_recode()`

```{r chapter_5}
diamonds=diamonds %>% as_tibble()

diamonds$cut=fct_recode(diamonds$cut, "meh"="Fair", "Wow"="Premium")

summary(diamonds$cut)
```

### Collapse factor levels

Extremely useful command for when you have infrequent cases in one factor and need to combine it with another.

Works by specifying a series of new level names, each of which contains the information from the old variables. Format is as follows:

```{r 5.2, eval=FALSE}

fct_collapse(dataset$variable,
			NewLevelA=c("OldLevel1","Oldlevel2"), # NewLevelA is the new variable that contains both variables 1 and 2
			NewLevelB=c("OldLevel3"))
```

### Add levels to a factor

use `fct_expand()`

```{r 5.3}
print("temp")
```

### Drop unused levels

Use `fct_drop()`

```{r 5.4}
print("temp")
```

### Change the order of a factor's levels

```{r 5.5}

example_data=tribble(~person, ~condition,
                     "bob", "25 years",
                     "jane", "5 years",
                     "jim", "5 years",
                     "john", "25 years")

example_data$condition=factor(example_data$condition)

str(example_data$condition)
```

Notice that R thinks these are nominal factors, and that 25 comes before 5. To fix this and correct the level order...

```{r chapter_5_end}
example_data$condition =fct_relevel(example_data$condition, c("5 years", "25 years")) # specify level order

str(example_data$condition)
```

