# Clean Data

## Replace a value with NA

Use `dplyr::na_if()` if you have a value coded in your data (e.g., 999) that you want to convert to NA

```{r 4.1}

example_data=dplyr::tribble(~name, ~bday_month,
                            "Ryan", 10,
                            "Z", 3,
                            "Jen", 999, 
                            "Tristin", 999,
                            "Cassidy", 6)

example_data

example_data$bday_month=na_if(example_data$bday_month, 999) #example doing one column at a time 

example_data

example_data %>% # can also pass the data to mutate and do it the tidyverse way
  mutate(bday_month=na_if(bday_month, 999)) 
```

## Replace NA's with a value

`tidyr::replace_na()` is very useful if you have some NA's in your data and you want to fill them in with some value.

```{r 4.2}
example_data=tibble::tribble(~name, ~fav_color, ~fav_food,
                             "Ryan", "green", "Mexican",
                             "Cassidy", "blue", NA,
                             "Z", NA, NA,
                             "Tristin", "purple", NA,
                             "Tarika", NA, NA,
                             "Jen", NA, "Italian")

example_data

# replace NA's in one col
tidyr::replace_na(example_data$fav_food, "MISSING")

# replace in multiple columns
example_data %>% mutate(across(c(fav_color, fav_food), replace_na, "MISSING"))
```

## Identify columns or rows with Missing values

`is.na()` is the base R way to identify, in a TRUE/FALSE manner, whether or not there are missing values in a vector

```{r 4.3}
y <- c(1,2,3,NA)

is.na(y) # returns a vector (F F F T)
```

## Find the percentage of a variable that is missing

Sometimes necessary to check before conducting an analysis. *This requires my package*, `legaldmlab`

```{r 4.4}
?legaldmlab::count_missing

mtcars %>% 
  select(hp:drat) %>% 
  legaldmlab::count_missing()
```

## Exclude Missing values from analysis

## Dropping Missing values from the data set

Use `tidyr::drop_na()` to remove rows with missing values.

```{r 4.5}
example_data=dplyr::tribble(~name, ~bday_month, ~car,
                            "Ryan", 10, "kia",
                            "Z", NA, "toyota",
                            "Jen", NA, NA,
                            "Tristin", 999, NA,
                            "Cassidy", 6, "honda")


knitr::kable(example_data)

example_data %>% drop_na() # with nothing specified, it drops ALL variables that have >=1 missing value

example_data %>% drop_na(car) # drops only rows with values missing in the specified column

```

### Filter out duplicates

```{r}

# Assuming your data frame is called 'my_data' and the column with duplicates is 'ID'

filtered_data <- my_data |> distinct(ID, .keep_all = TRUE)
```
