# Clean Data

```{r echo=false, message=FALSE, warning=FALSE}
library(tidyverse)
```


## Replace a value with NA

Use `dplyr::na_if()` if you have a value coded in your data (e.g., 999) that you want to convert to NA

```{r}

example_data=dplyr::tribble(~name, ~bday_month,
                            "Ryan", 10,
                            "Z", 3,
                            "Jen", 999, 
                            "Tristin", 999,
                            "Cassidy", 6)

example_data

example_data$bday_month=na_if(example_data$bday_month, 999) #example doing one column at a time 

example_data

example_data %>% # can also pass the data to mutate and do it the tidyverse way
  mutate(bday_month=na_if(bday_month, 999)) 
```

**Note.** `tidyr::replace_na()` does exactly the same thing. You can use either command.


## Identify columns or rows with Missing values

`is.na()` is the base R way to identify, in a TRUE/FALSE manner, whether or not there are missing values in a vector

```{r}
y <- c(1,2,3,NA)

is.na(y) # returns a vector (F F F T)
```

## Find the percentage of a variable that is missing

Sometimes necessary to check before conducting an analysis. *This requires my package*, `legaldmlab`

```{r}
?legaldmlab::count_missing

mtcars %>% 
  select(hp:drat) %>% 
  legaldmlab::count_missing()
```



## Exclude Missing values from analysis



# Dropping Missing values from the data set

`tidyr::drop_na()`

```{r}
example_data=dplyr::tribble(~name, ~bday_month, ~car,
                            "Ryan", 10, "kia",
                            "Z", NA, "toyota",
                            "Jen", NA, NA,
                            "Tristin", 999, NA,
                            "Cassidy", 6, "honda")


example_data

example_data %>% drop_na() # with nothing specified, it drops ALL variables that have >=1 missing value

example_data %>% drop_na(car) # drops only rows with values missing in the specified column

```



