# Working with Factors


## Manually recode/change a factor's levels

Use `forcats::fct_recode()`

```{r chapter_5}
diamonds=diamonds %>% as_tibble()

diamonds$cut=fct_recode(diamonds$cut, "meh"="Fair", "Wow"="Premium")

summary(diamonds$cut)
```

## Collapse factor levels

Extremely useful command for when you have infrequent cases in one factor and need to combine it with another.

Works by specifying a series of new level names, each of which contains the information from the old variables. Format is as follows:
```{r eval=FALSE}

fct_collapse(dataset$variable,
			NewLevelA=c("OldLevel1","Oldlevel2"), # NewLevelA is the new variable that contains both variables 1 and 2
			NewLevelB=c("OldLevel3"))
```
			

## Add levels to a factor

use `fct_expand()`

```{r}
print("temp")
```


## Drop unused levels

Use `fct_drop()`
```{r}
print("temp")
```


## Change the order of a factor's levels

```{r }

example_data=tribble(~person, ~condition,
                     "bob", "25 years",
                     "jane", "5 years",
                     "jim", "5 years",
                     "john", "25 years")

example_data$condition=factor(example_data$condition)

str(example_data$condition)
```

Notice that R thinks these are nominal factors, and that 25 comes before 5. To fix this and correct the level order...

```{r chapter_5_end}
example_data$condition =fct_relevel(example_data$condition, c("5 years", "25 years")) # specify level order

str(example_data$condition)
```


