# Working with Strings


## Remove a pattern from a string

Tidyverse command: `str_remove()`

```{r}
library(rvest)
library(tidyverse)

html=read_html('https://shop.tcgplayer.com/price-guide/pokemon/base-set') %>% 
  html_table(fill = TRUE)

html

# Saved as a list by default. Now extract your table from said list
html=as_tibble(html[[1]] %>% # find out which number it is in the list
                 select('PRODUCT','Rarity','Number','Market Price')) # if needed, specify which columns you want too

html

# BASE R METHOD (sub by replacing something with nothing)
gsub('\\$','',html$`Market Price`) #(what you're looking for, replace with, from where)

# TIDYVERSE METHOD (just remove)
str_remove(html$`Market Price`, pattern = "\\$")
```


## Replace one pattern in a string with another

Tidyverse command: `str_replace()`
Base R command: `gsub()`

```{r}

```


## Find (i.e., filter for) All instances of a string

Useful for finding very specific things inside a column (e.g., one particular person's name in a roster of names; everyone with a particular last name)

Tidyverse command: `str_detect()`
Base R command: `grepl()`

Note both must be nested inside of `filter()`
```{r eval=FALSE}

gradebook %>% filter(str_detect(student,"Rebecca"))

legaldmlab::pkmn %>% filter(str_detect(name, "Mewtwo"))
```

## Force all letters to lower case

Use `str_to_lower()`

```{r}
blah=tribble(~A, ~B,
             "A","X",
             "A","X")

blah

blah$A=str_to_lower(blah$A)

blah
```

