# Working with Strings


## Remove a pattern from a string

```{r chapter_6}

price_table=tribble(~car, ~price,
        "Corvette", "$65,000",
        "Mustang GT", "$40,000")

# BASE R METHOD (sub by replacing something with nothing)
gsub("\\$", "",price_table$price) # (pattern, replace with, object$column)

# TIDYVERSE METHOD
str_remove(price_table$price, pattern = "\\$")
```


## Replace one pattern in a string with another

Tidyverse command: `str_replace()`
Base R command: `gsub()`

```{r 6.2, eval=FALSE}
# base R
gsub(mtcars, replacement = )

#tidyverse


```


## Find (i.e., filter for) all instances of a string

Useful for finding very specific things inside a column (e.g., one particular person's name in a roster of names; everyone with a particular last name)

Tidyverse command: `str_detect()`
Base R command: `grepl()`

Note both must be nested inside of `filter()`
```{r 6.3, eval=FALSE}

cars_df=rownames_to_column(mtcars, var = "car")

# base R
cars_df |> filter(grepl("Firebird", car))

# tidyverse
cars_df %>% filter(str_detect(car,"Firebird"))
```

You can also search for **multiple strings simultaneously** by including the "or" logical operator *inside* the quotes.

```{r 6_filtermultiple, eval=FALSE}
cars_df |> filter(str_detect(car, "Firebird|Fiat"))
```

You can also include the negation logical operator to filter for all instances *except* those with the specified string.

```{r eval=FALSE}
# base R
cars_df |> filter(!(grepl("Pontiac", car)))

# tidyverse
cars_df |> filter(!(str_detect(car, "Pontiac")))
```

## Drop all rows from a data set that contain a certain string

```{r eval=FALSE}
# Tidyverse method
cars_df |> 
  filter(str_detect(car, "Merc", negate = TRUE)) #including negate=TRUE will negate all rows with the matched string

# base R
cars_df[!grepl("Merc", cars_df$car),]
```



## Force all letters to lower case

Use `stringr::str_to_lower()`

```{r 6.5, chapter_6_end}
blah=tribble(~A, ~B,
             "A","X",
             "A","X")

blah

blah$A=str_to_lower(blah$A)

blah
```

