# Figures and Graphs

There are three parts to a ggplot2 call: 1. data 2. aesthetic mapping 3. Layer

There is no piping involved in ggplot. You simply invoke ggplot, and tell it what they dataset is. Then you specify the aesthetics, and then the mapping. Lastly, include other optional stuff (e.g. expanded y-axis scale; titles and legends; etc.)

Every single plot has the exact same layout that ONLY USES the above three points:

```{r eval=FALSE}
ggplot(dataframe, aes(graph dimensions and variables used)) +
  geom_GraphType(specific graph controls)
```

Then if you have other stuff you want to add on top of this, like axis labels, annotations, highlights, etc., you keep adding those in separate lines

## Commands for graph types

| Graph Type    | Geom command       |
|---------------|--------------------|
| Scatter       | `geom_point()`     |
| Line          | `geom_line()`      |
| Box           | `geom_boxplot()`   |
| Bar           | `geom_bar()`       |
| Column        | `geom_col()`       |
| Histogram     | `geom_histogram()` |
| Density curve | `geom_density()`   |

## Quick notes for Histograms and Density Curves




## Highlight specific points

The `gghighlight` package is great for this

```{r}
ggplot(mtcars, aes(x= mpg, y=hp))+
  geom_point()+
  theme_classic()+
  ggrepel::geom_text_repel(data = mtcars, aes(label = hp))+ # add data labels (optional)
  gghighlight::gghighlight(hp > 200) # add highlights, according to some criteria


# example 2
diamonds_abr=diamonds %>% slice(1:100)

ggplot(diamonds_abr, aes(x= cut, y= price, colour=price))+
  geom_point()+
  theme_classic()+
  ggrepel::geom_text_repel(data = diamonds_abr, aes(label = price))+ # this line labels
  gghighlight::gghighlight(cut %in% c("Very Good", "Ideal")) #this line highlights

```


## Add labels to data points

```{r}
ggplot(mtcars, aes(x= mpg, y=hp))+
  geom_point()+
  theme_classic()+
  ggrepel::geom_text_repel(data = mtcars, aes(label = hp))

ggplot(mtcars, aes(x= mpg, y=hp))+
  geom_point() + geom_text(aes(label=hp, hjust=2.5, vjust=2.5))

geom_label(aes(label = scales::comma(n)), size = 2.5, nudge_y = 6)
```



## Plotting multiple graphs at once

## Change the graph/figure colors (bars; columns; dots; etc.)

### Adding color or fill

Nearly every single geom has color, fill, or both. **To add color**, simply add `color= variable_name` within the `aes()` command

```{r}
# add a color scale to the dots
ggplot(mtcars, aes(x= mpg, y=hp, color=mpg))+
  geom_point()

# change the fill of the bars
ggplot(mtcars, aes(factor(gear), fill=factor(carb)))+
  geom_bar()

# Change the outline of the bars by adding color inside the geom_bar() command
ggplot(mtcars, aes(factor(gear), fill=factor(carb)))+
  geom_bar(color="black")
```


### Changing the colors/fill

```{r}
ggplot(mtcars, aes(x= mpg, y=hp, color=hp))+
  geom_point()


ggplot(mtcars, aes(x = mpg, y = disp, colour = hp))+
  geom_point(size = 2) +
  scale_colour_got(option = "Lannister")

library(gameofthrones)

ggplot(mtcars, aes(factor(cyl), fill=factor(vs)))+
  geom_bar() +
  scale_fill_got(discrete = TRUE, option = "Daenerys")

ggplot(mtcars, aes(factor(gear), fill=factor(carb)))+
  geom_bar() +
  scale_fill_got(discrete = TRUE, option = "Tully")

ggplot(mtcars, aes(x = mpg, y = disp, colour = hp))+
  geom_point(size = 2) +
  scale_colour_got(option = "Lannister")


```

You can change colors to whatever you want with `scale_fill_manual()`. This command takes either a named color (e.g., "red"), or an RGB specification (\#RRGGBB), where each of the pairs RR,GG,BB consist of two hexadecimal digits giving a value in the range 00 to FF

Transparent is "\#RRGGBBAA"

```{r}
ggplot(mtcars, aes(factor(gear), fill=factor(carb)))+
  geom_bar() +
  scale_fill_manual(values=c("green", "yellow", "orange", "red", "purple", "blue"))
```

## Adding and Customizing Text

### Add a title, axis labels, and captions

Titles and axis labels are added via `labs()`.

```{r}
ggplot(mtcars, aes(x=cyl))+
    geom_bar(colour="gray", fill="lightgreen")+
  labs(title = "Ages of Survey Respondants by Group",
       x="Age Group",
       caption="Note. Younger= ages 11-29; Older= ages 30-86.")
```
### Center graph title

Add the line `theme(plot.title = element_text(hjust = 0.5))`

```{r}

ggplot(mtcars, aes(x=cyl))+
    geom_bar(colour="gray", fill="lightgreen")+
  labs(title = "Ages of Survey Respondants by Group",
       x="Age Group",
       caption="Note. Younger= ages 11-29; Older= ages 30-86.")+
  theme(plot.title = element_text(hjust = 0.5))
```


### Use different fonts

See tutorial on [this web page](https://www-r--bloggers-com.cdn.ampproject.org/v/s/www.r-bloggers.com/2021/07/using-different-fonts-with-ggplot2/amp/?amp_gsa=1&amp_js_v=a6&usqp=mq331AQIKAGwASCAAgM%3D#amp_tf=From%20%251%24s&aoh=16259111950507&csi=0&referrer=https%3A%2F%2Fwww.google.com&ampshare=https%3A%2F%2Fwww.r-bloggers.com%2F2021%2F07%2Fusing-different-fonts-with-ggplot2%2F)

Or, use the `extrafont` package, and set everything using the `theme()` command.

```{r}
# Visualize new groups
library(extrafont)
loadfonts(device="win")

ggplot(mtcars, aes(x=cyl))+
    geom_bar(colour="gray", fill="lightgreen")+
  labs(title = "Ages of Survey Respondants by Group",
       x="Age Group",
       caption="Note. Younger= ages 11-29; Older= ages 30-86.")+
  theme(plot.title = element_text(hjust = 0.5))
  theme(axis.title = element_text(face = "bold", family = "Courier New", size = 12),
        axis.text = element_text(face = "italic"),
        plot.caption = element_text(face = "italic", family = "Calibri", size = 9),
        plot.title = element_text(face = "bold",size = 14, family = "Courier New"))

```

## Remove gridlines

Add `theme(panel.grid = element_blank())`

```{r}
ggplot(mtcars, aes(x=cyl))+
    geom_bar(colour="gray", fill="lightgreen")+
  labs(title = "Ages of Survey Respondants by Group",
       x="Age Group",
       caption="Note. Younger= ages 11-29; Older= ages 30-86.")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title = element_text(face = "bold", family = "Courier New", size = 12),
        axis.text = element_text(face = "italic"),
        plot.caption = element_text(face = "italic", family = "Calibri", size = 9),
        plot.title = element_text(face = "bold",size = 14, family = "Courier New"))+
    theme(panel.grid = element_blank())+
```




## Faceting

This is dividing one plot into subplots, in order to communicate relationships better.
Again, this is just a single extra command, this time at the end of the code: facet_wrap(~columnhead)
The tilde sign in R means "by", as in "divide (something) by this"

```{r}
ggplot(gapminder_1952,aes(x=pop,y=lifeExp))+
  geom_point()+
  scale_x_log10()+
  facet_wrap(~continent)
```


This line produces a graph of population and life expectency, breaking it down to make a separate graph per each continent

## Log transformations

Sometimes when your data is really squished together on a graph it is hard to read. In this case, log transformations are really helpful, to change the scale of the data.

For example, by multiplying all your points by 10x

To create a log transformation of the same scatter plot above, add one extra bit: scale_x_log10()

```{r}
ggplot(gapminder_2007, aes(x= gdpPerCap, y= lifeExp))+ 
  geom_point() + 
  scale_x_log10()
```

You can also make both axis be logged by adding +scale again for y

## Changing the scale of the axis

Add `coord_cartesian(xlim = c(lower,upper))`

```{r}

```


## Add a regression line

Add the line `geom_smooth(method = "lm", formula = y ~ x)`

```{r}
ggplot(mtcars, aes(x= mpg, y=hp, color=mpg))+
  geom_point()+
  geom_smooth(method = "lm", formula = y ~ x)
```

