# Making Tables with flextable

NOTES: - **j** refers to the column - **i** refers to the row number

## APA Table Components

## Indent values

<https://davidgohel.github.io/flextable/reference/padding.html> <https://stackoverflow.com/questions/64134725/indentation-in-the-first-column-of-a-flextable-object>

Use the `padding` function:

```{r chapter_8, eval=FALSE}
ft <- padding(ft, i=2, j=1, padding.left=20)
```

## Add a Horizontal border (AKA horizontal "spanner")

```{r 8.2, eval=FALSE}
hline(., i=4, j=1:2, part = "body")
```

## Change font and font size

```{r 8.3, eval=FALSE}
  glm_table<-flextable::font(glm_table, part = "all", fontname = "Times") %>% # Font
                        fontsize(., size = 11, part = "all") # Font size
```

## Grouped table

```{r 8.4, eval=FALSE}

cars=rownames_to_column(mtcars, var = "Model")
test=flextable::as_grouped_data(x=cars, groups = c("cyl"))
```

## Adjust line spacing

```{r 8.5, eval=FALSE}
flextable::line_spacing(space = "0.5")
```

## Set global options for all flextables

```{r}
# set global options for all flextables
flextable::set_flextable_defaults(font.family = "Times New Roman", font.size = 11)

# create an fp_border object to set the borders for the flextables
border.test=officer::fp_border(color = "black", style = "solid", width = 1)
```

## Complete Example

```{r flextable_example_full}

library(flextable)
# set global options for all flextables
flextable::set_flextable_defaults(font.family = "Times New Roman", font.size = 11)

# create an fp_border object to set the borders for the flextables
border.test=officer::fp_border(color = "black", style = "solid", width = 1)


mtcars |> 
  slice(1:5) |> 
  flextable() |>
  # FIX BORDERS IN THE HEADER
  hline_bottom(border = border.test, part = "header") |>
  hline_top(border = border.test, part = "header") |>
  # CREATE A TITLE HEADER; APPLY FORMATTING
  add_header_lines(values = "Car Stuff") |> 
  hline_top(border = fp_border_default(width = 0), part = "header") |> 
  flextable::align(align = "left", part = "header", i=1) |> 
  italic(part = "header", i=1) |> 
  # FIX BORDER IN TABLE BODY
  hline_bottom(border = border.test, part = "body") |> 
  # Line spacing
  flextable::line_spacing(space = "0.5") |> 
  # Add a footer
  flextable::add_footer_lines(values = "A test footer") |> 
  # SET COLUMN WIDTH/DIMENSIONS
  autofit(part = "all") |> 
  width(j=1, width = 0.4, unit = "in")
```

## Wraper-function to create APA-style tables

This took way too many hours. You're welcome.

```{r}
APA_table=function(flextable_object, table_title, include_note){
  
  
  flextable_object=flextable_object |> 
    hline_bottom(border = border.test, part = "header") |>
    hline_top(border = border.test, part = "header") |>
    # CREATE A TITLE HEADER; APPLY FORMATTING
    add_header_lines(values = table_title) |> 
    hline_top(border = fp_border_default(width = 0), part = "header") |> 
    flextable::align(align = "left", part = "header", i=1) |> 
    italic(part = "header", i=1) |> 
    # FIX BORDER IN TABLE BODY
    hline_bottom(border = border.test, part = "body") |> 
    # SET FONT
    flextable::font(part = "all", fontname = "Times New Roman") |> 
    flextable::fontsize(part = "all", size = 11) |> 
    # SET COLUMN WIDTH/DIMENSIONS
    autofit(part = "all") 
  
  if(include_note==FALSE) (return(flextable_object))
  if(include_note!=FALSE) (flextable_object=add_footer_lines(flextable_object, 
                                                             values = paste0("Note."," ", include_note)) |> 
                             fontsize(part = "footer", size = 11) |>  
                             font(part = "footer", fontname = "Times"))
  
  return(flextable_object)
  
}
```
