<<<<<<< HEAD
# Functions, Loops, and Iterative Programming

## Functions

A function is a command that performs a specified operation and returns an output in accordance with that operation.

```{r eval=FALSE}
# example structure

Function_name=function(argument){
  Expressions
  return(output)
}
```

- *Argument* is your input. It is the thing you want to perform the operation on.
- *Expressions* is the actual operation (or operations) you want to perform on the supplied argument
- *return* tells R to return the result of the Expression to you when done.

**Example.** This function takes an input of numbers in the form of a vector and subtracts two from each.

```{r}
numbers=c(2,10,12,80)

sub_2=function(vector){
  result=vector-2
  return(result)
}

sub_2(numbers)
```

You can literally make a function to do anything you want!

***Limitations to your average, everyday functions***
However, there will be times when you want to perform some kind of function to multiple items, perhaps even simultaneously. For example, finding the mean of every column in a data set. To do this, you will probably find yourself repeating the same lines of code over and over, which violates a cardinal rule of efficient coding: Never paste the same code more than twice. If you have an iterative process or a task that needs to be repeated two or more times, consider how to replace the multiple lines of code with a For Loop or an iterative function.



## For-loops

A for loop is essentially  a function that applies a function or given set of operations iteratively on multiple things at once, and returns an output of many items.

For example, this code finds the means of every vector/column in a dataset by repeatedly applying the same code over and over to element "i" in the given list:

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

output <- vector("double", ncol(df))  # 1.Output. Create the object you want the results of the loop stored in.

for (i in seq_along(df)) {            # 2.Sequence of operations. "For each item 'i' along data frame…"
  output[[i]] <- median(df[[i]])      # 3.Body:"every individual item in 'output' = the median of each col in df
}

output
```

Check out [this book chapter](https://r4ds.had.co.nz/iteration.html#mapping-over-multiple-arguments) for a great and detailed explanation of for-loops and functional coding

Although for loops are nice, they can be somewhat unwieldy. R programmers typically use iterating functions instead! 

Examples of iterating functions are the lapply, vapply, sapply, etc. family of base R commands. But these can be confusing and the commands are not great.

The purrr package offers a better way to do iterating functions over base R; it's the tidyverse way to make efficient and understandable for loops! **If you have a need for a for-loop for something, see the next section instead on how to use purrr to make an iterative function. Important to understand conceptually what a for-loop is, but using them is impractical when you have purrr**

## Purrr and iterative loops
*All notes here come from Charlotte Wickham's lecture tutorial [on YouTube](https://www.youtube.com/watch?v=7UlWJWfZO9M)

Purrr's map functions offer a way to apply any existing function (even functions you've made) to multiple things at once (be it lists; data frame columns, individual items in vector, etc.).

When you have tasks that you want to do repeatedly, you can either copy and paste your code a bunch of times, or make an iterative function that does each task for you. You could use a for-loop to do these tasks, but {purrr} is a package that lets you build iterating functions, which have advantages over for-loops.



### Useful ways to use map()

#### Apply a single function to every column in a data frame

Pipe the whole data frame to the map command:
```{r}
df %>% map_df()
```


#### Apply a function to every item in a column/vector
Pipe the data frame and column/variable, and include the [ ] to index by vector position:
```{r}
BF_output$Evidence= BF_output$BF[] %>% map_chr(effectsize::interpret_bf)
```


#### Import several files in a directory at once (batch importing)

**Note.** This reads many .csv files into a SINGLE, unified data frame.
If you want to import many files at once but keep them separated, you'll need a different command

```{r}
# first run these commands
data_dir <-"log dump"
fs::dir_ls(data_dir)
csv_files <- fs::dir_ls(data_dir, regexp = "\\.csv$")

# then run this
File_name_here=csv_files %>% 
  map_dfr(read_csv)
```



### Walkthrough Case Example

This is an example walk through showing how we can use `purrr` to speed things up dramatically and/or reduce the use of unwanted, extra code in our scripts. In this guide I'll be building a table of LPGA Tour statistics from multiple webpages.

If I tried to do each table individually, I'd be copying and pasting and re-writing the same code up to **four times** to get each table downloaded, extracted, and formatted the way I want before I could unify them with `left_join`. By using `purrr::map()`, we can cut our needed code down by 75%.

Not only does this mean we have a tidier script, but also, things will be less messy, and we have less chances for error (since we are coding less). And of course, we don't have the annoyances of redoing multiple sections of code.

```{r}
pacman::p_load(rvest, tidyverse)
```

The workflow for purrr goes like this:

First, you want to figure out how to do each step of your process line-by-line, for a single item. The idea is to try and walk through each step of the process and see exactly what will need to be done each each step and what the code will like, *before trying to code it all at once at a higher level*.

Once you have each step for the first item figured out, make functions for each step that condense that code down to one command.

Lastly, apply each function from your individual steps to all items in your list by using `purr::map()`.


### Do for One

#### 1. Scrape data

```{r}
# line
html1=read_html("https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=04") %>% 
  html_table(fill = TRUE)


# function
quick_read_html=function(url){
  url=url %>% read_html(url) %>%
    html_table(fill = TRUE) 
}

html=quick_read_html("https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=04")
```


#### 2. Find out which element of the list the desired table is in

```{r}
as_tibble(test[[2]])
```

#### 3. Extract the table

```{r}
#line
table1=as_tibble(test[[2]])


# function version
extract_tibble=function(html_list){
  df=as_tibble(html_list[[2]])
  return(df)
}

table1=extract_tibble(html1)
```


#### 4. Format table

```{r}
#line
table1=table1 %>% 
  select(-X1) %>% 
  rename("Name"="X2") %>% 
  slice(2:n())

head(table1)


# function
test_format=function(table){
  table=table %>% 
    select(-X1) %>% 
    rename("Name"="X2") %>% 
    slice(2:n())
  
  return(table)
}

test_format(table1)
```

### DO FOR ALL

```{r}

#### FUNCTIONS ####
# Read
quick_read_html=function(url){
  url=url %>% read_html(url) %>%
    html_table(fill = TRUE) 
}

# Extract
extract_tibble=function(html_list){
  df=as_tibble(html_list[[2]])
  return(df)
}

# Format
test_format=function(table){
  table=table %>% 
    select(-X1) %>% 
    rename("Name"="X2") %>% 
    slice(2:n())
  
  return(table)
}

#### GO FOR IT ####

URLs=c("https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=04",
       "https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=08",
       "https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=06",
       "https://scores.nbcsports.com/golf/averages.asp?tour=LPGA&rank=12")


tables_list=URLs %>% 
  map(quick_read_html) %>% #read and scrape every URL
  map(extract_tibble) %>% # Get rid of the extra list levels
  map(test_format) # Apply any formatting that would be universal to all of them at once


lpga_data= tables_list %>% 
  reduce(left_join, by="Name") %>% # Combine all list levels into a single tibble, matching by the "Name" column
  rename("Avg_Drive"="X3.x", "Avg_Putts"="X3.y", # final, non-universal formatting
         "Rounds_Played"="X3.y.y" , "Avg_Score"="X4", 
         "Greens_Hit"="X3.x.x")

# VOILA! 
head(lpga_data)

```

