# Advanced Coding Tips

## Grammar and Syntax

### Regex expressions and symbols

```{r chapter_13, eval=FALSE}
str_remove(html$`Market Price`, pattern = "$") # doesn't remove the $ sign
str_remove(html$`Market Price`, pattern = "\\$") # works
```

### The colon-equals (`:=`) operator

Sometimes when making a function you need to use the colon-equals operator, rather than just the normal \<- or = assignment operators

Specifically, when you have multiple named arguments in your function...Read my question and someone's answer on this blogpost: <https://community.rstudio.com/t/help-creating-simple-function/109011/2>

```{r 13.1, eval=FALSE}
function(df, col, newCol_name){
  
  # drop all NA's so the function can work properly
  df_NAdropped=df |> drop_na({{col}})
  
  # apply the function, saving it in a small tibble with only the outlier column and a key to join by 
  outliers_key=df_NAdropped |> 
    dplyr::mutate(outliers=dplyr::if_else(round(abs({{col}}-median({{col}}))/(1.483*mad({{col}}, constant = 1)),2)>2.24,1,0)) |> 
    select(subject_id, outliers)
  
  # join back together
  df=df |> left_join(outliers_key, by=c("subject_id")) 
  
  num.outliers=df |> 
    dplyr::filter(outliers==1) |> 
    dplyr::count()
  
  df$outliers=replace_na(df$outliers, 999)
  df=df |> rename({{newCol_name}}:="outliers")
  
  message(paste0(num.outliers, " outlier(s) detected"))
  return(df)
}
```

### User-supplied expressions or named columns in functions

This is when you have to put double braces around something...

```{r}

```

### When a command requires a named column or data set, but you've already supplied it and it's required a second time

If you're writing a function with a pipe but the command you're using needs the data set defined in it, you specify it as `.x` , or simply just `.`; Here is an example:

```{r}

```

## Creating a package

<https://rstudio4edu.github.io/rstudio4edu-book/data-pkg.html>

### Documenting package meta-data

<https://r-pkgs.org/description.html>

### Connecting to other packages

<https://kbroman.org/pkg_primer/pages/depends.html>

### Linking Git and Github

[view this detailed guide](https://r-pkgs.org/git.html#git-rstudio) by Jenny Bryan, and [this YouTube video](https://www.youtube.com/watch?v=XbSwdHBiO4E) if you want the full guide; or just follow the TL;DR below.

**Quick summary of steps in YouTube video**:

1.  Open project folder in Windows Explorer and click in the URL bar, then type `cmd` to open command prompt

2.  If there are any pre-existing git files or repository info there, remove it with the following: `rd .git /S/Q`

3.  Tell git to create a new repo by typing: `git init`

4.  Then tell it to include all files in the current place by typing: `git add .`

5.  Commit these files with: `git commit -m "Initial commit"`

6.  At this point you've created a git and GitHub repo each; now link them with: `git remote add origin [https URL of GitHub repo]`

7.  Push all these changes live with: `git push -u origin master`

## Creating a bookdown

<https://www.youtube.com/watch?app=desktop&v=m5D-yoH416Y&feature=youtu.be>

### Rendering the book once it's done

1.  Render locally with bookdown::render_book("index.Rmd")

2.  Use `browseURL("docs/index.html")` to view your book locally (or just open `index.html` in a browser).

3.  If it looks good, commit and push all changed files to GitHub.
